---
title: "Homework week 1"
author: "Viviana Romero Alarocn"
format: html
editor: visual
---

Source : [https://github.com/alarconvv/UmassClasses/tree/master/Biol607](Github%20Repository)

## Let's explore the dataset quakes looking at earthquakes off of Fiji.

1.  Load quakes with `data(quakes)`. Show what's there with str() and summary(). What do you see?

```{r}
# loading data
data("quakes")

# see data
head(quakes)

# see structure
str(quakes)

# see summary
summary(quakes)
```

2.  Show the entirety of the column `long`

```{r}
# display "long" column

quakes$long
```

3.  Hey, another useful function - unique()! Apply it to a vector, and you can see what are all of the unique values. It's great for really digging into a problematic vector.

```{r}
# see the unique values of station
unique(quakes$stations)
```

4.  What unique stations are there? Use `length()` with `unique()` to determine how many stations there are.

```{r}
# # see how many stations there are
length(unique(quakes$stations))
```

5.  Using `range()`, what is the range of depths where quakes occur?

```{r}
# ranges of depths where the quakes occurred
range(quakes$depth)
```

6.  Impress Yourself. Where was the earthquake of largest magnitude found? You'll need to use come comparisons and the `max()` function for this in your indices instead of a number!

```{r}
# see the largest magnitude quakes

max(quakes$mag)

quakes$stations[which(quakes$mag == max(quakes$mag))]



```

Rta: the biggest earthquake, with a magnitude of 6.4, was recorded in the station 122

------------------------------------------------------------------------

Meta 1. For some of you, this week was your first experience with R (or even programming at all) - how did it feel to connect concepts to a novel set of problems. For those who have experience with R, was this a cakewalk or challenging? How did it feel to shake off the cobwebs?

RTa: I have worked with R for long time, however it is good to remember basic things.

Meta 2. How much time did this take you, roughly?

Rta: 15 minutes

Meta 3. Please give yourself a weak/sufficient/strong assessment on this assignment. Feel free to comment on why.

Rta: strong. I have good bases in R but every time that I come back to the basic things, I can discover new ways to resolve the same problem. So, I am happy in this class.
