---
title: "Loading Data"
author: "Viviana Romero"
format: html
editor: visual
---

## 

### Starting Data

dplyr is a package for helping with tabular data manipulation.

```{r}
#libraries
library(tidyverse)

# download data


download.file(url = "https://ndownloader.figshare.com/files/2292169",destfile = "portal_data_joined.csv")
```

### Reading the data into R

```{r}
surveys <- read_csv("portal_data_joined.csv")
```

by default read_csv asssume commas as delimitators. if the data are delimited by ; you should use read_cvs2. read_tsv for dataset with \tab and read_delim with other deimitators.

we can see the basic structure using head

```{r}
head(surveys)
```

to open the data set into Rstudio View() could work

```{r}
View(surveys)
```

### What is a data frame

using tidyverse, whe we load data the object is stored in a object class tibble. data frame is a representation of data in form of table.

every column has the sameclass

. a easy way to see the structure is using str()

```{r}
str(surveys)
```

Exercise:

Based on the output of `str(surveys)`, can you answer the following questions?

-   What is the class of the object `surveys`?

    ```         
    tibble
    ```

-   How many rows and how many columns are in this object?

```         
786 × 13
```

### **Indexing and subsetting data frames**

in order to extract data from tibble

```{r}
# We can extract specific values by specifying row and column indices
# in the format: 
# data_frame[row_index, column_index]
# For instance, to extract the first row and column from surveys:
surveys[1, 1]

# First row, sixth column:
surveys[1, 6]   

# We can also use shortcuts to select a number of rows or columns at once
# To select all columns, leave the column index blank
# For instance, to select all columns for the first row:
surveys[1, ]

# The same shortcut works for rows --
# To select the first column across all rows:
surveys[, 1]

# An even shorter way to select first column across all rows:
surveys[1] # No comma! 

# To select multiple rows or columns, use vectors!
# To select the first three rows of the 5th and 6th column
surveys[c(1, 2, 3), c(5, 6)] 

# We can use the : operator to create those vectors for us:
surveys[1:3, 5:6] 

# This is equivalent to head_surveys <- head(surveys)
head_surveys <- surveys[1:6, ]

# As we've seen, when working with tibbles 
# subsetting with single square brackets ("[]") always returns a data frame.
# If you want a vector, use double square brackets ("[[]]")

# For instance, to get the first column as a vector:
surveys[[1]]

# To get the first value in our data frame:
surveys[[1, 1]]
```

See that : is a special fucntion to create vectors of integers and - is for excluding columns

```{r}
surveys[, -1]                 # The whole data frame, except the first column
surveys[-(7:nrow(surveys)), ] # Equivalent to head(surveys)

```

Data frames can be subset by calling indices, but also by calling their column names directly

```{r}
# As before, using single brackets returns a data frame:
surveys["species_id"]
surveys[, "species_id"]

# Double brackets returns a vector:
surveys[["species_id"]]

# We can also use the $ operator with column names instead of double brackets
# This returns a vector:
surveys$species_id
```

Exercise:

1.  Create a `data.frame` (`surveys_200`) containing only the data in row 200 of the `surveys` dataset.

    ```{r}
    surveys200 <- data.frame(surveys[200,])
    ```

2.  Notice how `nrow()` gave you the number of rows in a `data.frame`?

    ```{r}
    nrow(surveys)
    ```

-   Use that number to pull out just that last row from the `surveys` dataset.

    ```{r}
    surveys["34778",]
    ```

-   Compare that with what you see as the last row using `tail()` to make sure it's meeting expectations.

    ```{r}
    tail(surveys)
    ```

-   Pull out that last row using `nrow()` instead of the row number.

    ```{r}
    surveys[nrow(surveys),]
    ```

-   Create a new data frame (`surveys_last`) from that last row.

    ```{r}
    surveys_last <- data.frame(surveys[nrow(surveys),])
    ```

3.  Use `nrow()` to extract the row that is in the middle of the data frame. Store the content of this row in an object named `surveys_middle`.

    ```{r}
    surveys_middle <- data.frame(surveys[nrow(surveys)/2,])
    ```

4.  Combine `nrow()` with the `-` notation above to reproduce the behavior of `head(surveys)`, keeping just the first through 6th rows of the surveys dataset.

    ```{r}
    surveys[-(7:nrow(surveys)),]
    ```

### Factors

Integers usually are factors/ categorical data with levels

```{r}
surveys$sex <- factor(surveys$sex)

```

To order the levels in a factor column

```{r}
#order
sex <- factor(c("male", "female", "female", "male"), levels = c('males', 'female'))
# levels
levels(sex)
#nlevels
nlevels(sex)


sex
```

Exercise

1.  Change the columns `taxa` and `genus` in the `surveys` data frame into a factor.

    ```{r}
    surveys$taxa <- as.factor(surveys$taxa)

    surveys$genus <- as.factor(surveys$genus)
    ```

2.  How many rabbits were observed?

    ```{r}
    length(surveys$taxa == "Rabbit")
    ```

3.  How many different genera are in the `genus` column?

```{r}
nlevels(surveys$genus)
```

### Converting factors

we can switch of class using as.\*

```{r}
as.character(sex)
```

some classes has more priority than other.

```{r}
year_fct <- factor(c(1990, 1983, 1977, 1998, 1990))
as.numeric(year_fct)               # Wrong! And there is no warning...
as.numeric(as.character(year_fct)) # Works...
as.numeric(levels(year_fct))[year_fct] 

```

Renaming factors

factor are important for plots

```{r}
plot(surveys$sex)
```

```{r}
summary(surveys$sex)

levels(surveys$sex)


```

even when sex column has NAs it is not a factor, so we can add NA as part of the data

```{r}
sex <- surveys$sex

levels(sex)

sex <- addNA(sex)

levels(sex)

head(sex)

levels(sex)[3]

levels(sex)


```

Exercise

-   Rename \"F\" and \"M\" to \"female\" and \"male\" respectively.

    ```{r}
    levels(sex) <- c("female", "male", NA)
    ```

-   Now that we have renamed the factor level to \"undetermined\", can you recreate the barplot such that \"undetermined\" is first (before \"female\")?

```{r}
sex <- surveys$sex
levels(sex)

sex <- addNA(sex)
levels(sex)

head(sex)

levels(sex)[3] <- "undetermined"
levels(sex)

plot(sex)
```
