---
title: "Homework week 2"
author: "Viviana Romero Alarcon"
format: html
editor: visual
---

## QQPLOT

libraries

```{r}
library(tidyverse)
library(dplyr)
library(readr)
library(ggplot2)
library(forcats)

#pacman::p_load(tidyverse)

```

Set theme for ggplot

```{r}
theme_set(theme_bw(base_size=12))


```

1.  Read data

    ```{r}
    ice <- read_csv("http://biol607.github.io/homework/data/NH_seaice_extent_monthly_1978_2016.csv") %>%
      mutate(Month_Name = factor(Month_Name),
             Month_Name = fct_reorder(Month_Name, Month))
    ```

2.  Boxplot

Make a boxplot showing the variability in sea ice extent every month. IYKYK make it fancy.

```{r}
ggplot(data = ice, mapping = aes( x =Extent,  y = Month_Name))+
  geom_boxplot(color = 'grey50')+
  geom_point(aes(color = Month_Name), size= 0.7, alpha= 0.5)+
  stat_summary(color = 'darkred', shape = 18 )

```

3.  Ridgelines

Use ggridges to do the same thing. What do you learn that is different than the boxplot?

```{r}
library(ggridges)
library(ggplot2)

ggplot(data = ice, mapping = aes( x =Extent,  y = Month_Name)) +
  geom_density_ridges(alpha = 0.5,mapping = aes(color = Month_Name, fill = Month_Name ))

```
4. Making Groups from the Continuous
One thing that’s really cool about faceting is that you can use cut_*() functions on continuous variables to make facets by groups of continuous variables. To see what I mean, try cut_interval(1:10, n = 5) See how it makes five bins of even width? We use cut_interval() or other cut functions with faceting like so facet_wrap(vars(cut_interval(some_variable))).

```{r}
ggplot(data = ice, mapping = aes( x =Extent)) +
  geom_histogram() +
  facet_wrap(vars(Month_Name))
```


With the original data, plot sea ice by year, with different lines (oh! What geom will you need for that?) for different months. Then, use facet_wrap and cut_interval(Month, n=4) to split the plot into seasons.



